Urban and socio-economic correlates of property prices in Dublin's area

### 2020-08-20 manuscript/manuscript.Rmd
final submission

keywords: ["Property price", "Housing Market", "Feature Analysis", "Machine Learning", "Geocoding"]
abstract: |
  Understanding the characteristics of the housing market is essential for both sellers and buyers. However, the housing market is influenced by multiple factors. In this paper, the urban and socio-economic structure of an area is used to predict the price of 10387 properties sold in 2018 in the city of Dublin. More precisely, the direct distance from each property to 160 urban features taken from OpenStreetMap is calculated, and an extreme gradient boosting linear regression performed. Using these features, the model explains 45\% of the housing price variance. The most important features in this model are the proximity to an embassy and to a grassland. In addition, the results of a population census from 2016 are also used to correlate with the price of properties. From this census, 48 features are used as the input of a gradient boost linear regression model. In all, the socio-economic features are explaining 43\% of the housing price variance as well. The density of individuals reporting that they are not providing unpaid personal help for a friend or family member as well as individuals reporting that they have no religion are the most important socio-economic feratures. By taking into account either urban or socio-economic features, it is possible to accurately estimate housing prices and to predict their evolution.

>>>>

keywords: ["Property Price", "Housing Market", "Feature Analysis", "Machine Learning", "Geocoding"]
abstract: |
  Understanding the characteristics of the housing market is essential for both sellers and buyers. However, the housing market is influenced by multiple factors. In this paper, the urban and socio-economic structure of an area is used to predict the price of 10387 properties sold in 2018 in the city of Dublin. More precisely, the direct distance from each property to 160 urban features taken from OpenStreetMap is calculated, and an extreme gradient boosting linear regression performed. Using these features, the model explains 45\% of the housing price variance. The most important features in this model are the proximity to an embassy and to a grassland. In addition, the results of a population census from 2016 are also used to correlate with the price of properties. From this census, 48 features are used as the input of a gradient boost linear regression model. In all, the socio-economic features are explaining 43\% of the housing price variance as well. The density of individuals reporting that they are not providing unpaid personal help for a friend or family member as well as individuals reporting that they have no religion are the most important socio-economic feratures. By taking into account either urban or socio-economic features, it is possible to accurately estimate housing prices and to predict their evolution.

### 2020-08-20 manuscript/manuscript.tex
final submission

\hypersetup{
            pdftitle={Urban and socio-economic correlates of property prices in Dublin's area},
            pdfkeywords={Property price, Housing Market, Feature Analysis, Machine Learning, Geocoding},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls

@@

% keywords
\begin{IEEEkeywords}
Property price; Housing Market; Feature Analysis; Machine Learning; Geocoding
\end{IEEEkeywords}

@@

\caption{\label{tab:osm-model-comparison}Accuracy of XGBoost models using urban features according the type of gradient boosting method and output objective.}
\centering
\fontsize{8}{10}\selectfont
\begin{tabular}[t]{llrl}
\toprule
Objective & $RMSE$ & $R^2$ & $MAE$\\
\midrule
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{Tree based}}\\
\hspace{1em}squarederror & €127,783 & 0.44 & €85,115\\
\hspace{1em}squaredlogerror & €426,989 & - & €391,670\\
\hspace{1em}gamma & €129,052 & 0.43 & €87,039\\
\hspace{1em}tweedie & €126,626 & 0.45 & €85,073\\
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{Linear regression based}}\\
\hspace{1em}squarederror & €150,925 & 0.21 & €111,810\\
\hspace{1em}squaredlogerror & €406,947 & 0.00 & €369,709\\
\hspace{1em}gamma & €183,987 & 0.05 & €138,978\\
\hspace{1em}tweedie & €175,950 & 0.07 & €132,838\\
\bottomrule
\multicolumn{4}{l}{\textit{Note: }}\\
\multicolumn{4}{l}{Missing value ares due to incompatibilities between data and methods.}\\
\end{tabular}
\end{table}

@@

\caption{\label{tab:census-model-comparison}Accuracy of XGBoost models using socio-economic features according the type of gradient boosting method and output objective.}
\centering
\fontsize{8}{10}\selectfont
\begin{tabular}[t]{llrl}
\toprule
Objective & $RMSE$ & $R^2$ & $MAE$\\
\midrule
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{Tree based}}\\
\hspace{1em}squarederror & €131,270 & 0.42 & €86,892\\
\hspace{1em}squaredlogerror & €426,989 & - & €391,670\\
\hspace{1em}gamma & €129,221 & 0.43 & €86,949\\
\hspace{1em}tweedie & €130,927 & 0.42 & €86,707\\
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{Linear regression based}}\\
\hspace{1em}squarederror & €146,997 & 0.25 & €108,172\\
\hspace{1em}squaredlogerror & €406,126 & 0.01 & €368,916\\
\hspace{1em}gamma & €185,183 & 0.08 & €138,255\\
\hspace{1em}tweedie & €173,179 & 0.10 & €129,087\\
\bottomrule
\multicolumn{4}{l}{\textit{Note: }}\\
\multicolumn{4}{l}{Missing value ares due to incompatibilities between data and methods.}\\
\end{tabular}
\end{table}

@@

The authors would like to thank the developers of the following R packages used to process, analyze, display and report data: R (Version 4.0.2; R Core Team 2020) and the R-packages \emph{dplyr} (Version 1.0.1; Wickham, François, et al. 2020), \emph{forcats} (Version 0.5.0; Wickham 2020), \emph{ggplot2} (Version 3.3.2; Wickham 2016), \emph{here} (Version 0.1; Müller 2017), \emph{kableExtra} (Version 1.1.0; Zhu 2019), \emph{latex2exp} (Version 0.4.0; Meschiari 2015), \emph{lubridate} (Version 1.7.9; Grolemund and Wickham 2011), \emph{lwgeom} (Version 0.2.5; Pebesma 2020), \emph{magrittr} (Version 1.5; Bache and Wickham 2014), \emph{mgcv} (Version 1.8.31; Wood 2011, 2004, 2003; Wood et al. 2016), \emph{nlme} (Version 3.1.148; Pinheiro et al. 2020), \emph{OpenStreetMap} (Version 0.3.4; Fellows and JMapViewer library by Jan Peter Stotz 2019), \emph{osmdata} (Version 0.1.3; Padgham et al. 2017), \emph{papaja} (Version 0.1.0.9997; Aust and Barth 2020), \emph{patchwork} (Version 1.0.1; Pedersen 2019), \emph{purrr} (Version 0.3.4; Henry and Wickham 2020), \emph{readr} (Version 1.3.1; Wickham, Hester, and Francois 2018), \emph{sf} (Version 0.9.5; Pebesma 2018), \emph{stringr} (Version 1.4.0; Wickham 2019), \emph{tibble} (Version 3.0.3; Müller and Wickham 2020), \emph{tidyr} (Version 1.1.0; Wickham and Henry 2020), \emph{tidyverse} (Version 1.3.0; Wickham, Averick, et al. 2019), \emph{tmap} (Version 3.1; Tennekes 2018), \emph{xgboost} (Version 1.1.1.1; Chen et al. 2020), and \emph{yardstick} (Version 0.0.7; Kuhn and Vaughan 2020).

>>>>

\hypersetup{
            pdftitle={Urban and socio-economic correlates of property prices in Dublin's area},
            pdfkeywords={Property Price, Housing Market, Feature Analysis, Machine Learning, Geocoding},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls

@@

% keywords
\begin{IEEEkeywords}
Property Price; Housing Market; Feature Analysis; Machine Learning; Geocoding
\end{IEEEkeywords}

@@

\caption{\label{tab:osm-model-comparison}Accuracy of XGBoost models using urban features according the type of gradient boosting method and output objective.}
\centering
\fontsize{8}{10}\selectfont
\begin{tabular}[t]{llrl}
\toprule
Objective & $RMSE$ & $R^2$ & $MAE$\\
\midrule
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{Tree based}}\\
\hspace{1em}squarederror & €127,783 & 0.44 & €85,115\\
\hspace{1em}squaredlogerror & €426,989 & - & €391,670\\
\hspace{1em}gamma & €129,052 & 0.43 & €87,039\\
\hspace{1em}tweedie & €126,626 & 0.45 & €85,073\\
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{Linear regression based}}\\
\hspace{1em}squarederror & €150,847 & 0.21 & €111,786\\
\hspace{1em}squaredlogerror & €406,939 & 0.00 & €369,708\\
\hspace{1em}gamma & €177,238 & 0.07 & €133,973\\
\hspace{1em}tweedie & €171,480 & 0.08 & €129,138\\
\bottomrule
\multicolumn{4}{l}{\textit{Note: }}\\
\multicolumn{4}{l}{Missing value ares due to incompatibilities between data and methods.}\\
\end{tabular}
\end{table}

@@

\caption{\label{tab:census-model-comparison}Accuracy of XGBoost models using socio-economic features according the type of gradient boosting method and output objective.}
\centering
\fontsize{8}{10}\selectfont
\begin{tabular}[t]{llrl}
\toprule
Objective & $RMSE$ & $R^2$ & $MAE$\\
\midrule
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{Tree based}}\\
\hspace{1em}squarederror & €131,270 & 0.42 & €86,892\\
\hspace{1em}squaredlogerror & €426,989 & - & €391,670\\
\hspace{1em}gamma & €129,221 & 0.43 & €86,949\\
\hspace{1em}tweedie & €130,927 & 0.42 & €86,707\\
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{Linear regression based}}\\
\hspace{1em}squarederror & €146,963 & 0.25 & €108,151\\
\hspace{1em}squaredlogerror & €406,213 & 0.02 & €369,023\\
\hspace{1em}gamma & €183,008 & 0.08 & €136,445\\
\hspace{1em}tweedie & €174,731 & 0.10 & €130,483\\
\bottomrule
\multicolumn{4}{l}{\textit{Note: }}\\
\multicolumn{4}{l}{Missing value ares due to incompatibilities between data and methods.}\\
\end{tabular}
\end{table}

@@

The authors would like to thank the developers of the following R packages used to process, analyze, display and report data: R (Version 4.0.2; R Core Team 2020) and the R-packages \emph{dplyr} (Version 1.0.2; Wickham, François, et al. 2020), \emph{forcats} (Version 0.5.0; Wickham 2020), \emph{ggplot2} (Version 3.3.2; Wickham 2016), \emph{here} (Version 0.1; Müller 2017), \emph{kableExtra} (Version 1.1.0; Zhu 2019), \emph{latex2exp} (Version 0.4.0; Meschiari 2015), \emph{lubridate} (Version 1.7.9; Grolemund and Wickham 2011), \emph{lwgeom} (Version 0.2.5; Pebesma 2020), \emph{magrittr} (Version 1.5; Bache and Wickham 2014), \emph{mgcv} (Version 1.8.31; Wood 2011, 2004, 2003; Wood et al. 2016), \emph{nlme} (Version 3.1.148; Pinheiro et al. 2020), \emph{OpenStreetMap} (Version 0.3.4; Fellows and JMapViewer library by Jan Peter Stotz 2019), \emph{osmdata} (Version 0.1.3; Padgham et al. 2017), \emph{papaja} (Version 0.1.0.9997; Aust and Barth 2020), \emph{patchwork} (Version 1.0.1; Pedersen 2019), \emph{purrr} (Version 0.3.4; Henry and Wickham 2020), \emph{readr} (Version 1.3.1; Wickham, Hester, and Francois 2018), \emph{sf} (Version 0.9.5; Pebesma 2018), \emph{stringr} (Version 1.4.0; Wickham 2019), \emph{tibble} (Version 3.0.3; Müller and Wickham 2020), \emph{tidyr} (Version 1.1.1; Wickham and Henry 2020), \emph{tidyverse} (Version 1.3.0; Wickham, Averick, et al. 2019), \emph{tmap} (Version 3.1; Tennekes 2018), \emph{xgboost} (Version 1.1.1.1; Chen et al. 2020), and \emph{yardstick} (Version 0.0.7; Kuhn and Vaughan 2020).
